Combination of field and experimental data together with computational models

### 2022-02-22 manuscript_1.0/manuscript_1.0.Rmd
Error found in the manual coding of the cleaner entries used to compute visitor preference. In loadFieldData.R a programatic way of coding cleaner entries was done. Now that the manual and programatic results coincide I can re-run the chains. New figure configurations were implemented. Among those supplementary figures showing the diagnostics of the MCMC. A new result section was written.

abstract: |
  While it is generally straightforward to quantify individual performance in cognitive experiments, identifying the underlying cognitive processes remains a major challenge. Often, different mechanistic underpinnings yield similar performances, and Lloyd Morgan’s cannon warrants acceptance of the simpler explanation. Alternatively, when the different mechanisms interact with environmental conditions, variation in performance across environments might allow to statistically infer the mechanism responsible. However, to do this it is necessary to have quantitative predictions for the different candidate mechanisms. Here, we use a set of computational models to get quantitative predictions from alternative learning mechanisms, and using Bayesian statistics fit the model parameters to performance data. We used experimental data on performance in an ephemeral reward task by wild-caught cleaner fish *Labroides dimidiatus*, as well as cleaner and client fish densities from the locations of capture. The task can in principle be solved by estimating future consequences of an action, or by perceiving the removal of the ephemeral reward as psychological punishment (negative reinforcement). We found that a model where cleaners only estimate the future consequences of their actions explains best which cleaner relative abundances cause the fish to develop a preference for an ephemeral food source. This model also yields performances that can be considered the result of locally optimal decision-rules, in contrast to the negative reinforcement model.  We argue that the combination of computational models with data is a powerful tool to infer the mechanistic underpinning behind cognitive performance.

@@

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,warning = FALSE,message=FALSE)
library(here)
```

@@

When researchers aimed at testing wild-caught cleaner fish, as 
well as individuals from other species, abilities 
to prioritize visitor over resident clients in a lab-based paradigm, 
they used Plexiglas plates of different colours and shapes offering same 
amount of food as surrogates for visitor and resident clients.
One plate is made to behave like a resident, i.e. it remains until 
the cleaner had eaten off it. The other plate was made to behave 
like a visitor, i.e. it is removed if not inspected first.
Cleaners learned to prefer the visitor plate and
hence obtained the double amount of food [@bshary_Asymmetric_2002].
Furthermore, adult cleaner fish outperformed various primates as well as rats and
pigeons in this original version of the biological market task
[@salwiczek_Adult_2012; @zentall_Early_2017]; yet the African grey
parrots solve this task as well [@pepperberg_Can_2014]. Given that primates 
are known to readily distinguish 
between one and two reward units if presented in more conventional ways 
$\color{red}{\text{ref}}$, this raises the question:
what makes the task difficult for most non-cleaners?

@@

<!-- A further model analysis contrast the effect of these two  -->
<!-- parameters against variation in the relative abundance of cleaners  -->
<!-- and the two client types. Assuming certainty that visitors leave, the  -->
<!-- two parameters yield different  -->
<!-- predictions when relative client abundance is very high. Under such  -->
<!-- conditions, the analysis shows that under very high client (relative)  -->
<!-- abundance the presence of future rewards ($\gamma>0$) no longer favours  -->
<!-- the preference for the ephemeral clients, while the preference persists  -->
<!-- when the negative reward mediated by client behaviour ($\eta>0$) is  -->
<!-- included. Thus, the analysis suggests that future reward as an  -->
<!-- explanation for cleaner ability to prefer visitors is supported.  -->
<!-- At first sight, empirical data seem to support future rewards as  -->
<!-- the mechanism used by cleaners, as the statistical analysis  by  -->
<!-- Triki *et al.* [@triki_Biological_2019] data revealed that low  -->
<!-- cleaner-to-client ratios predict low preference for visitors and hence  -->
<!-- low performance in the ephemeral reward task. However, low cleaner-to-client  -->
<!-- ratios may also lead to low probabilities of visitors leaving, which would  -->
<!-- cause low preferences for visitors independently of the underlying  -->
<!-- mechanism [@quinones_Reinforcement_2019]. Furthermore, a combination  -->
<!-- of future and negative rewards, could still be yield a better explanation  -->
<!-- for the data than any single factor.  -->

@@

<!-- Cleaner densities, in interaction with visitor client leaving probability,  -->
<!-- explained best individual performance in the ephemeral reward task  -->
<!-- [@triki_Biological_2019], while cleaner-to-client ratio was not a significant  -->
<!-- factor. This creates a certain mismatch with the model by Quiñones *et al.*  -->
<!-- [@quinones_Reinforcement_2019], which focused on abundance of cleaners relative  -->
<!-- to overall client abundance in order to capture the characteristics  -->
<!-- of the market. Importantly, relative abundances of clients define not only  -->
<!-- the probability of interactions with residents and visitors, but also affect  -->
<!-- how often the cleaning station is empty (e.g. there are no clients to be cleaned).  -->
<!-- The higher the relative abundance of cleaners is, the higher is the probability  -->
<!-- of cleaning stations being empty at a given time point. How often the  -->
<!-- cleaning stations are empty in turns influences the value that cleaners can  -->
<!-- expect from their choices. The emptier the cleaning stations are, the more  -->
<!-- beneficial it is to prioritize a client that will not stay. On the other hand,  -->
<!-- situations where cleaners have to choose will be rare if cleaner relative  -->
<!-- abundance is high, which reduces the total value of prioritizing visitors  -->
<!-- over residents and makes it more difficult to learn a preference for visitors.  -->
<!-- Given the model structure, we used the empirical data to calculate relative  -->
<!-- abundances of cleaners, residents and visitors for the current analyses.  -->
<!-- To incorporate the empirical finding that absolute densities matter, we  -->
<!-- introduced a new scaling constant (see section on model fitting).  -->

@@

The cognitive performance data was from a total of 120 cleaners 
(10 individuals per 12 site/year) tested in the 
ephemeral reward task [@triki_Biological_2019;@triki_Brain_2020]. Authors 
housed all captured cleaners individually in glass aquaria 
( $62cm \times 27cm \times 37 cm$ ) and provided them 
with PVC pipes ($10 cm \times 1 cm$) as shelters. 
The task consisted of exposing the cleaner fish to substitute 
models of client fish in the form of two *Plexiglas* plates offering the 
same amount of food (one item of mashed prawn).  The two plates differed 
in colour and pattern (horizontal green stripes or vertical pink stripes) 
but had equal size ($10 cm \times 7 cm$). Importantly, the two plates played 
different roles as either a visitor (ephemeral food source) or 
resident (permanent food source). That is, if a cleaner inspected the 
resident plate first, the experimenter withdrew the visitor plate out of 
the aquarium as a consequence.  Choosing first the visitor plate, 
however, granted access to both plates. The equal size of the plates 
forced cleaner fish to learn to give service priority to the visitor plate 
based solely on the behaviour-cue of the plates rather than size-cue 
[@wismer_Cuebased_2019]. Triki *et al.* [@triki_Biological_2019;@triki_Brain_2020] 
tested the fish for a maximum of 200 trials with 20 trials a day, 10 trials 
in the morning and 10 trials in the afternoon. They randomized and 
counterbalanced the plates’ spatial location (i.e. left or right) 
between trials. Similarly, they counterbalanced the plates’ decoration 
(colour and pattern) and the plates’ role (visitor or resident) between the 
tested fish. In the original studies, once a fish reach a learning criterion, 
that is, performing significantly above chance level ($> 50%$, $p-value ≥ 0.05$),
they passed to a reversal version of the task where the roles of the 
visitor/resident Plexiglas plates were swapped (see Triki et al. 2019, 2020). 
Here, we used instead a subset of these data in order to have an idea on 
cleaner fish preferences for the visitor plate, even if they do not reach 
the learning criterion within 200 trials. To do so, we first extracted the 
trial-by-trial outcomes from the last two session (20 trials) of those who 
never reached the learning criterion for visitor plate 
$\color{red}{\text{(N = XX cleaner fish)}}$. 
For those who reached the learning criterion at some point during the 
test and passed to a reversal phase, we extracted the trials-by-trial 
outcomes from the last session (10 trials) before passing to reversal 
and the last 10 trials they were exposed to in the test 
$\color{red}{\text{(N = XX cleaner fish)}}$.

@@

To capture with the model the relationship between the ecological 
variables and cleaner fish preferences for visitors we need to scale 
the absolute population densities of cleaner fish from the empirical data
to a measure of relative abundances that captures 
client visitation patterns. This is because in the model, relative 
abundances of clients define not only the probability of residents and visitors, 
but also how often the cleaning station is empty (e.g. there are no 
clients to be cleaned). The frequency with which clients visit the station 
is another variable, which in nature may vary among different client species 
depending on their ectoparasite loads. We do not have field estimates 
for species-specific parasite loads, especially not as a function of site. 
In order to control for these aspects, we computed a measure of relative 
cleaner fish abundance for each reef site relative to 
absolute abundances and multiplied it
by a scaling constant that changes the range of the variable. This scaling 
constant is hence meant to capture variation in the market conditions driven 
partly by cleaner abundance. We fitted the value of the scaling constant as 
part of the statistical inference. As for the visitor and resident 
abundances we computed a relative measure with respect to the 
total client abundance, and weighted that by the rescaled relative absence 
of cleaners ($1-\text{relative cleaner abundance}$). Thus, all three measures 
of relative abundance sum up to 1, and can be used in the model as a proxy 
for the probability of having different options in the cleaning station. 
Note that we introduced the scaling constant to control for variation that 
is not captured by the model; its parameter distribution does not offer 
biological insights.

Once we calculated the relative abundances, we obtained predictions from 
the model for each one of the locations and ran the Markov Chain. We started the
chain with random values for the parameters of interest, then ran the 
computational model once for every reef site using as input the ecological 
explanatory variables. The model outputs the probability of choosing the visitor
for each location $p_i$, where $i$ is the index for the 12 locations. Assuming a 
binomial distribution, the probability that each of the cleaner fish 20 choices 
in the ephemeral reward task was generated by the model 
is given by $\binom{n}{k_ij}p^{k_{ij}}_i (1-p_i)^{n-k_{ij}}$;
where $k_{ij}$ is the number of times that cleaner fish $j$ in reef site $i$ 
chose the visitor over the resident; and $n=20$ (due to our choice of using 20 
choices per cleaner fish). By taking the natural logarithm and summing over all 
the cleaner fish and reef sites we obtained the log-likelihood of the data 
given the model and parameters. We then proposed a new set of parameters drawn 
from a uniform distribution centred around the old parameter set. The amplitude 
of the uniform distribution used for each parameter can be found in Table 
\ref{tab:param}. Subsequently, we ran the model and calculated the likelihood 
with the new parameter set. We then used the ratio of the two likelihoods to 
choose which parameter set to keep. New parameter sets with a higher likelihood 
than the old set replaced old ones, and those with a lower likelihood replaced 
current ones  with a probability equals to the log-likelihood ratio. 
Given that we only used the likelihoods in the decision, we used a 
uninformative prior. Once we decided whether the new parameter set replaces 
the old one, we ran the model again to sample the likelihood 
distribution of the parameter set. We then started 
the cycle again by proposing a new set of parameters, and repeated
the process for $1e^5$ steps. We ran 5 independent chains, 
discarded the first $100$ samples of each chain as burn in, 
and after that we kept 1 in every 100 samples to avoid autocorrelation. 
The collection of parameters sets kept in 
all chains approximates the posterior distribution. We coded the model 
as well as the fitting algorithm in *c++*, the diagnostics and visualization 
in R [@rcoreteam_Language_2021]. All codes are accessible at 
$\color{red}{\text{include citation}}$

@@

```{r post, fig.cap="Posterior distributions for parameter values $\\gamma$, $\\eta$ and the scaling constant. We show the kernel density estimates, below the mode (black dot) and the 65\\% (light blue shade) and 95\\% (grey shade)  highest posterior density interval for the three paramteres. On the left, panels a, b and c show posterior distributions for a model with all three parameters; while on the right, d and e, show a model with only chaining. Panel f shows a measure of fit for both models, namely the distribution of pseudo-$R^2$ obtained from sampling the posterior distribution of parameter values.",out.width='100%',fig.height=6}
library(data.table)
library(coda)
library(ggplot2)
library(cowplot)
library(ggdist)
source(here("../R_files/posPlots.R"))
source(here("loadData.R"))
library("jsonlite")

@@

# NLables names for the three scenarios
labels.Scen<-c("both","gam","Nrew")
labelsPlot.Scen<-c("FM","C","OC")

@@

MCMCdata1<-loadMCMCrun(scen1,thinning = thinning,burn.in = burn.in)
MCMCdata2<-loadMCMCrun(scen2,thinning = thinning,burn.in = burn.in)
MCMCdata3<-loadMCMCrun(scen3,thinning = thinning,burn.in = burn.in)

@@

mcmcList.2<-mcmc.list(do.call(list,lapply(unique(MCMCdata2$seed), function(repli){
  rundata<-MCMCdata2[seed==repli]
  mcmcRun<-mcmc(rundata[,.(gamma,scaleConst)],thin = thinning)#
  return(mcmcRun)
})))

mcmcList.2<-mcmc.list(do.call(list,lapply(unique(MCMCdata2$seed), function(repli){
  rundata<-MCMCdata2[seed==repli]
  mcmcRun<-mcmc(rundata[,.(negReward,scaleConst)],thin = thinning)#
  return(mcmcRun)
})))

mcmcList.3<-mcmc.list(do.call(list,lapply(unique(MCMCdata3$seed), function(repli){
  rundata<-MCMCdata3[seed==repli]
  mcmcRun<-mcmc(rundata[,.(gamma,negReward,scaleConst)],thin = 100)#
  return(mcmcRun)
})))

@@

cuts.2<-lapply(MCMCdata2[,.(gamma,scaleConst)],
               function(x){
                 c(mode_hdi(x,.width = intervals[c(2,1)])$ymin,
                   mode_hdi(x,.width = intervals)$ymax)  
               })

cuts.3<-lapply(MCMCdata3[,.(negReward,scaleConst)],
               function(x){
                 c(mode_hdi(x,.width = intervals[c(2,1)])$ymin,
                   mode_hdi(x,.width = intervals)$ymax)  
               })

# get all the likelihoods from the posterior for all models
loglikehoods.all<-data.table(lglikelihood=c(MCMCdata1$fit,MCMCdata2$fit,
                                             MCMCdata3$fit),
                              model=c(rep(labels.Scen[1],length(MCMCdata1$fit)),
                                      rep(labels.Scen[2],length(MCMCdata2$fit)),
                                      rep(labels.Scen[3],length(MCMCdata3$fit))
                                      ))

# Read field data to calculate null log-likelihood
fieldData<-fread(here("Data","data_ABC_cleaner_abs.txt"))

nullLikehood<-sum(dbinom(x=fieldData[,score_visitor],
                         size = 20,prob = 0.5,log = TRUE))

# calculate pseudo R^2
loglikehoods.all[,Rsqrd:=1-lglikelihood/nullLikehood]

panel.rsqr.all<-
  ggplot(data=loglikehoods.all[is.finite(lglikelihood)],
                  aes(y=model,x=Rsqrd,fill=model))+
  stat_halfeye(alpha=0.5,point_size=3)+
  theme_classic()+xlim(-0.5,0.25)+
  scale_fill_manual(values=c("#1b9e77","#d95f02","#7570b3"),
  name = "Model", labels = labelsPlot.Scen)+
  scale_y_discrete(labels=labelsPlot.Scen)+
  geom_vline(xintercept = 0,color="black",size=0.5)+
  theme(legend.position = 'none')
          # c(.32, .65)
        # ,legend.key.size = unit(0.22,'cm'),
        # axis.text.y = element_blank(),
        # legend.text = element_text(size=7))

gam.both.post<-ggplot(data=MCMCdata1,aes(x=gamma)) + 
  stat_halfeye(aes(fill=stat(cut(x,breaks = cuts.1$gamma))),
               point_interval = mode_hdi, .width = c(.66, .95),
               point_size=3,show.legend=FALSE) + 
  labs(title=labelsPlot.Scen[1],
       subtitle = expression(gamma),x="",y="")+
  theme_classic()+
  scale_fill_manual(values=c("gray85","skyblue","gray85"))
nrew.both.post<-ggplot(data=MCMCdata1,aes(x=negReward)) + 
  stat_halfeye(aes(fill=stat(cut(x,breaks = cuts.1$negReward))),
               point_interval = mode_hdi, 
               .width = c(.66, .95),point_size=3,
               show.legend=FALSE)+theme_classic()+
  scale_fill_manual(values=c("gray85","skyblue","gray85"))+
  labs(title="",subtitle = expression(eta),x="",y="")+
  xlim(-2,2)
gam.gam.post<-ggplot(data=MCMCdata2,aes(x=gamma)) +
  stat_halfeye(aes(fill=stat(cut(x,breaks = cuts.2$gamma))),
               point_interval = mode_hdi, .width = c(.66, .95),
               point_size=3,
               show.legend=FALSE) +
  scale_fill_manual(values=c("gray85","skyblue","gray85"))+
  labs(title=labelsPlot.Scen[2],
       subtitle = expression(gamma),x="",y="")+
  theme_classic()
nrew.nrew.post<-ggplot(data=MCMCdata3,aes(x=negReward)) +
  stat_halfeye(aes(fill=stat(cut(x,breaks = cuts.3$negReward))),
               point_interval = mode_hdi, .width = c(.66, .95),
               point_size=3,
               show.legend=FALSE) +
  scale_fill_manual(values=c("gray85","skyblue","gray85"))+
  labs(title=labelsPlot.Scen[3],
       subtitle = expression(gamma),x="",y="")+
  theme_classic()+xlim(-2,2)

first.col<-plot_grid(nrow=3,align = "v",byrow = TRUE,
          gam.both.post,gam.gam.post,nrew.nrew.post,           
          labels=c('a','c','d'))
second.col<-plot_grid(nrow = 2,align = "v",byrow = TRUE,
          nrew.both.post,rsqr.both,rel_heights = c(1,2),
          labels = c('b','e'))
plot_grid(ncol = 2,first.col,second.col)

@@

>>>>

abstract: |
  While it is generally straightforward to quantify individual performance in cognitive experiments, identifying the underlying cognitive processes remains a major challenge. Often, different mechanistic underpinnings yield similar performances, and Lloyd Morgan’s cannon warrants acceptance of the simpler explanation. Alternatively, when the different mechanisms interact with environmental conditions, variation in performance across environments might allow to statistically infer the mechanism responsible. However, to do this it is necessary to have quantitative predictions for the different candidate mechanisms. Here, we use a set of computational models to get quantitative predictions from alternative learning mechanisms, and using Bayesian statistics fit the model parameters to performance data. We used experimental data on performance in an ephemeral reward task by wild-caught cleaner fish *Labroides dimidiatus*, as well as cleaner and client fish densities from the locations of capture. The task can in principle be solved by estimating future consequences of an action, or by perceiving the removal of the ephemeral reward as psychological punishment (negative reinforcement). We found that a model where cleaners only estimate the future consequences of their actions explains best which cleaner fish relative abundances cause the fish to develop a preference for an ephemeral food source. This model also yields performances that can be considered the result of locally optimal decision-rules, in contrast to the negative reinforcement model.  We argue that the combination of computational models with data is a powerful tool to infer the mechanistic underpinning behind cognitive performance.

@@

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,warning = FALSE,message=FALSE)
library(here)
source(file=here("loadFieldData.R"))
```

@@

When researchers aimed at testing wild-caught cleaner fish, as 
well as individuals from other species, abilities 
to prioritize visitor over resident clients in a lab-based paradigm, 
they used Plexiglas plates of different colours and shapes offering same 
amount of food as surrogates for visitor and resident clients.
One plate is made to behave like a resident, i.e. it remains until 
the cleaner fish had eaten off it. The other plate was made to behave 
like a visitor, i.e. it is removed if not inspected first.
Cleaners learned to prefer the visitor plate and
hence obtained the double amount of food [@bshary_Asymmetric_2002].
Furthermore, adult cleaner fish outperformed various primates as well as rats and
pigeons in this original version of the biological market task
[@salwiczek_Adult_2012; @zentall_Early_2017]; yet the African grey
parrots solve this task as well [@pepperberg_Can_2014]. Given that primates 
are known to readily distinguish 
between one and two reward units if presented in more conventional ways 
$\color{red}{\text{ref}}$, this raises the question:
what makes the task difficult for most non-cleaners?

@@

<!-- A further model analysis contrast the effect of these two  -->
<!-- parameters against variation in the relative abundance of cleaners  -->
<!-- and the two client types. Assuming certainty that visitors leave, the  -->
<!-- two parameters yield different  -->
<!-- predictions when relative client abundance is very high. Under such  -->
<!-- conditions, the analysis shows that under very high client (relative)  -->
<!-- abundance the presence of future rewards ($\gamma>0$) no longer favours  -->
<!-- the preference for the ephemeral clients, while the preference persists  -->
<!-- when the negative reward mediated by client behaviour ($\eta>0$) is  -->
<!-- included. Thus, the analysis suggests that future reward as an  -->
<!-- explanation for cleaner fish ability to prefer visitors is supported.  -->
<!-- At first sight, empirical data seem to support future rewards as  -->
<!-- the mechanism used by cleaners, as the statistical analysis  by  -->
<!-- Triki *et al.* [@triki_Biological_2019] data revealed that low  -->
<!-- cleaner-to-client ratios predict low preference for visitors and hence  -->
<!-- low performance in the ephemeral reward task. However, low cleaner-to-client  -->
<!-- ratios may also lead to low probabilities of visitors leaving, which would  -->
<!-- cause low preferences for visitors independently of the underlying  -->
<!-- mechanism [@quinones_Reinforcement_2019]. Furthermore, a combination  -->
<!-- of future and negative rewards, could still be yield a better explanation  -->
<!-- for the data than any single factor.  -->

@@

<!-- Cleaner densities, in interaction with visitor client leaving probability,  -->
<!-- explained best individual performance in the ephemeral reward task  -->
<!-- [@triki_Biological_2019], while cleaner-to-client ratio was not a significant  -->
<!-- factor. This creates a certain mismatch with the model by Quiñones *et al.*  -->
<!-- [@quinones_Reinforcement_2019], which focused on abundance of cleaners relative  -->
<!-- to overall client abundance in order to capture the characteristics  -->
<!-- of the market. Importantly, relative abundances of clients define not only  -->
<!-- the probability of interactions with residents and visitors, but also affect  -->
<!-- how often the cleaning station is empty (e.g. there are no clients to be cleaned).  -->
<!-- The higher the relative abundance of cleaners is, the higher is the probability  -->
<!-- of cleaning stations being empty at a given time point. How often the  -->
<!-- cleaning stations are empty in turns influences the value that cleaners can  -->
<!-- expect from their choices. The emptier the cleaning stations are, the more  -->
<!-- beneficial it is to prioritize a client that will not stay. On the other hand,  -->
<!-- situations where cleaners have to choose will be rare if cleaner fidh relative  -->
<!-- abundance is high, which reduces the total value of prioritizing visitors  -->
<!-- over residents and makes it more difficult to learn a preference for visitors.  -->
<!-- Given the model structure, we used the empirical data to calculate relative  -->
<!-- abundances of cleaners, residents and visitors for the current analyses.  -->
<!-- To incorporate the empirical finding that absolute densities matter, we  -->
<!-- introduced a new scaling constant (see section on model fitting).  -->

@@

The cognitive performance data was from a total of 120 cleaners 
(10 individuals per 12 site/year) tested in the 
ephemeral reward task [@triki_Biological_2019;@triki_Brain_2020]. Authors 
housed all captured cleaners individually in glass aquaria 
( $62cm \times 27cm \times 37 cm$ ) and provided them 
with PVC pipes ($10 cm \times 1 cm$) as shelters. 
The task consisted of exposing the cleaner fish to substitute 
models of client fish in the form of two *Plexiglas* plates offering the 
same amount of food (one item of mashed prawn).  The two plates differed 
in colour and pattern (horizontal green stripes or vertical pink stripes) 
but had equal size ($10 cm \times 7 cm$). Importantly, the two plates played 
different roles as either a visitor (ephemeral food source) or 
resident (permanent food source). That is, if a cleaner fish inspected the 
resident plate first, the experimenter withdrew the visitor plate out of 
the aquarium as a consequence.  Choosing first the visitor plate, 
however, granted access to both plates. The equal size of the plates 
forced cleaner fish to learn to give service priority to the visitor plate 
based solely on the behaviour-cue of the plates rather than size-cue 
[@wismer_Cuebased_2019]. Triki *et al.* [@triki_Biological_2019;@triki_Brain_2020] 
tested the fish for a maximum of 200 trials with 20 trials a day, 10 trials 
in the morning and 10 trials in the afternoon. They randomized and 
counterbalanced the plates’ spatial location (i.e. left or right) 
between trials. Similarly, they counterbalanced the plates’ decoration 
(colour and pattern) and the plates’ role (visitor or resident) between the 
tested fish. In the original studies, once a fish reach a learning criterion, 
that is, performing significantly above chance level ($> 50%$, $p-value ≥ 0.05$),
they passed to a reversal version of the task where the roles of the 
visitor/resident Plexiglas plates were swapped (see Triki et al. 2019, 2020). 
Here, we used instead a subset of these data in order to have an idea on 
cleaner fish preferences for the visitor plate, even if they do not reach 
the learning criterion within 200 trials. To do so, we first extracted the 
trial-by-trial outcomes from the last two session (20 trials) of those who 
never reached the learning criterion for visitor plate 
($N = `r fieldData.cleaner.filt.sum[V1==2,length(V1)]`$ cleaner fish). 
For those who reached the learning criterion at some point during the 
test and passed to a reversal phase, we extracted the trials-by-trial 
outcomes from the last session (10 trials) before passing to reversal 
and the last 10 trials they were exposed to in the test 
($N = `r fieldData.cleaner.filt.sum[V1>2,length(V1)]`$ cleaner fish).

@@

To capture with the model the relationship between the ecological 
variables and cleaner fish preferences for visitors we need to scale 
the absolute population densities of cleaner fish from the empirical data
to a measure of relative abundances that captures client visitation patterns. 
This is because in the model, relative abundances of clients 
define not only the probability of residents and visitors, 
but also how often the cleaning station is empty (e.g. there are no 
clients to be cleaned). The frequency with which clients visit the station 
is another variable, which in nature may vary among different client species 
depending on their ectoparasite loads. We do not have field estimates 
for species-specific parasite loads, especially not as a function of site. 
In order to control for these aspects, we computed a measure of relative 
cleaner fish abundance for each reef site relative to 
absolute abundances and multiplied it
by a scaling constant that changes the range of the variable. This scaling 
constant is hence meant to capture variation in the market conditions driven 
partly by cleaner fish abundance. We fitted the value of the scaling constant as 
part of the statistical inference. As for the visitor and resident 
abundances we computed a relative measure with respect to the 
total client abundance, and weighted that by the rescaled relative absence 
of cleaners ($1-\text{relative cleaner abundance}$). Thus, all three measures 
of relative abundance sum up to 1, and can be used in the model as a proxy 
for the probability of having different options in the cleaning station. 
Note that we introduced the scaling constant to control for variation that 
is not captured by the model; its parameter distribution does not offer 
biological insights.

Once we calculated the relative abundances, we obtained predictions from 
the model for each one of the locations and ran the Markov Chain. We started the
chain with random values for the parameters of interest, then ran the 
computational model once for every reef site using as input the ecological 
explanatory variables. The model outputs the probability of choosing the visitor
for each location $p_i$, where $i$ is the index for the 12 locations. Assuming a 
binomial distribution, the probability that each of the cleaner fish 20 choices 
in the ephemeral reward task was generated by the model 
is given by $\binom{n}{k_ij}p^{k_{ij}}_i (1-p_i)^{n-k_{ij}}$;
where $k_{ij}$ is the number of times that cleaner fish $j$ in reef site $i$ 
chose the visitor over the resident; and $n=20$ (due to our choice of using 20 
choices per cleaner fish). By taking the natural logarithm and summing over all 
the cleaner fish and reef sites we obtained the log-likelihood of the data 
given the model and parameters. We then proposed a new set of parameters drawn 
from a uniform distribution centred around the old parameter set. The amplitude 
of the uniform distribution used for each parameter can be found in Table 
\ref{tab:param}. Subsequently, we ran the model and calculated the likelihood 
with the new parameter set. We then used the ratio of the two likelihoods to 
choose which parameter set to keep. New parameter sets with a higher likelihood 
than the old set replaced old ones, and those with a lower likelihood replaced 
current ones  with a probability equals to the log-likelihood ratio. 
Given that we only used the likelihoods in the decision, we used a 
uninformative prior. Once we decided whether the new parameter set replaces 
the old one, we ran the model again to sample the likelihood 
distribution of the parameter set. We then started 
the cycle again by proposing a new set of parameters, and repeated
the process for $1e^5$ steps. We ran 5 independent chains, 
discarded the first $1000$ samples of each chain as burn in, 
and after that we kept 1 in every 100 samples to avoid autocorrelation. 
The collection of parameters sets kept in 
all chains approximates the posterior distribution. We coded the model 
as well as the fitting algorithm in *c++*, the diagnostics and visualization 
in R [@rcoreteam_Language_2021]. All codes are accessible at 
$\color{red}{\text{include citation}}$

@@

(ref:post) Posterior distributions for parameter values $\gamma$ and $\eta$ for 
the three model ("Full model", "chaining only" and "penalty only"). We show 
the kernel density estimates, below the mode (black dot) and the 65% 
(light blue shade) and 95% (grey shade)  highest posterior density interval 
for the two parameters. On the top, panels a and b show posterior 
distributions for a full model, including chaining ($\gamma$) and penalty 
($\eta$). Panel c, shows the $\gamma$ estimate from model with only chaining. 
Panel d shows the $\\eta$ estimate from a model with only penalty. Panel e, 
shows a measure of fit for all models, namely the distribution of pseudo-$R^2$ 
obtained from sampling the posterior distribution of parameter values.

```{r post, fig.cap='(ref:post)',out.width='100%',fig.height=6}
library(data.table)
library(coda)
library(ggplot2)
library(cowplot)
library(ggdist)
library(ggpubr)
source(here("../R_files/posPlots.R"))
source(here("loadData.R"))
library("jsonlite")

@@

# NLables names for the three scenarios
labels.Scen<-c("both","gam","Nrew")
labelsPlot.Scen<-c("Full Model","Chaining","Penalty")

@@

MCMCdata1<-loadMCMCrun(scen1,thinning = thinning,burn.in = burn.in)
MCMCdata2<-loadMCMCrun(scen2,thinning = thinning,burn.in = burn.in)
# MCMCdata3<-loadMCMCrun(scen3,thinning = thinning,burn.in = burn.in)

@@

mcmcList.2<-mcmc.list(do.call(list,lapply(unique(MCMCdata2$seed), function(repli){
  rundata<-MCMCdata2[seed==repli]
  mcmcRun<-mcmc(rundata[,.(gamma,scaleConst)],thin = thinning)
  return(mcmcRun)
})))

# mcmcList.3<-mcmc.list(do.call(list,lapply(unique(MCMCdata3$seed), function(repli){
#   rundata<-MCMCdata3[seed==repli]
#   mcmcRun<-mcmc(rundata[,.(gamma,scaleConst)],thin = thinning)#
#   return(mcmcRun)
# })))

@@

cuts.2<-lapply(MCMCdata2[,.(gamma,scaleConst)],
               function(x){
                 c(mode_hdci(x,.width = intervals[c(2,1)])$ymin,
                   mode_hdci(x,.width = intervals)$ymax)
               })

# cuts.3<-lapply(MCMCdata3[,.(negReward,scaleConst)],
#                function(x){
#                  c(mode_hdci(x,.width = intervals[c(2,1)])$ymin,
#                    mode_hdci(x,.width = intervals)$ymax)
#                })

# get all the likelihoods from the posterior for all models
# loglikehoods.all<-data.table(lglikelihood=c(MCMCdata1$fit,MCMCdata2$fit,
#                                              MCMCdata3$fit),
#                               model=c(rep(labels.Scen[1],length(MCMCdata1$fit)),
#                                       rep(labels.Scen[2],length(MCMCdata2$fit)),
#                                       rep(labels.Scen[3],length(MCMCdata3$fit))
#                                       ))
# 
# # Read field data to calculate null log-likelihood
# fieldData<-fread(here("Data","data_cleaner_abs_threa1.5.txt"))
# 
# nullLikehood<-sum(dbinom(x=fieldData[,score_visitor],
#                          size = 20,prob = 0.5,log = TRUE))
# 
# # calculate pseudo R^2
# loglikehoods.all[,Rsqrd:=1-lglikelihood/nullLikehood]
# 
# 
# panel.rsqr.all<-
#   ggplot(data=loglikehoods.all[is.finite(lglikelihood)],
#                   aes(y=model,x=Rsqrd,fill=model))+
#   stat_halfeye(alpha=0.5,point_size=3)+
#   theme_classic()+#xlim(-0.5,0.25)+
#   scale_fill_manual(values=c("#1b9e77","#d95f02","#7570b3"),
#   name = "Model", labels = labelsPlot.Scen)+
#   scale_y_discrete(labels=labelsPlot.Scen)+
#   geom_vline(xintercept = 0,color="black",size=0.5)+
#   theme(legend.position = 'none')
#           # c(.32, .65)
#         # ,legend.key.size = unit(0.22,'cm'),
#         # axis.text.y = element_blank(),
#         # legend.text = element_text(size=7))
# 
# gam.both.post<-ggplot(data=MCMCdata1,aes(x=gamma)) +
#   stat_halfeye(aes(fill=stat(cut(x,breaks = cuts.1$gamma))),
#                point_interval = mode_hdi, .width = c(.66, .95),
#                point_size=3,show.legend=FALSE) +
#   labs(title=labelsPlot.Scen[1],
#        subtitle = expression(gamma),x="",y="")+
#   theme_classic()+
#   scale_fill_manual(values=c("gray85","skyblue","gray85"))
# 
# nrew.both.post<-ggplot(data=MCMCdata1,aes(x=negReward)) +
#   stat_halfeye(aes(fill=stat(cut(x,breaks = cuts.1$negReward))),
#                point_interval = mode_hdi,
#                .width = c(.66, .95),point_size=3,
#                show.legend=FALSE)+theme_classic()+
#   scale_fill_manual(values=c("gray85","skyblue","gray85"))+
#   labs(title="",subtitle = expression(eta),x="",y="")+
#   xlim(-2,2)
# gam.gam.post<-ggplot(data=MCMCdata2,aes(x=gamma)) +
#   stat_halfeye(aes(fill=stat(cut(x,breaks = cuts.2$gamma))),
#                point_interval = mode_hdi, .width = c(.66, .95),
#                point_size=3,
#                show.legend=FALSE) +
#   scale_fill_manual(values=c("gray85","skyblue","gray85"))+
#   labs(title=labelsPlot.Scen[2],
#        subtitle = expression(gamma),x="",y="")+
#   theme_classic()
# nrew.nrew.post<-ggplot(data=MCMCdata3,aes(x=negReward)) +
#   stat_halfeye(aes(fill=stat(cut(x,breaks = cuts.3$negReward))),
#                point_interval = mode_hdi, .width = c(.66, .95),
#                point_size=3,
#                show.legend=FALSE) +
#   scale_fill_manual(values=c("gray85","skyblue","gray85"))+
#   labs(title=labelsPlot.Scen[3],
#        subtitle = expression(gamma),x="",y="")+
#   theme_classic()+xlim(-2,2)
# 
# first.col<-plot_grid(nrow=3,align = "v",byrow = TRUE,
#           gam.both.post,gam.gam.post,nrew.nrew.post,
#           labels=c('a','c','d'))
# second.col<-plot_grid(nrow = 2,align = "v",byrow = TRUE,
#           nrew.both.post,panel.rsqr.all,rel_heights = c(1,2),
#           labels = c('b','e'))
# plot_grid(ncol = 2,first.col,second.col)

@@

(ref:diagnosticsfull) MCMC convergence diagnostics for the full model. 
On the left trace-plots, on the right changes along the chain of the 
Gelman and Rubin's shrink factor [@brooks_General_1998].

```{r diagnosticsfull, fig.cap='(ref:diagnosticsfull)',echo=FALSE,fig.height=5}

longMCMCdata1<-melt(MCMCdata1,id.vars = c("iteration","seed"),
                    variable.name = "parameter",
                    measure.vars = c("gamma","negReward","scaleConst"))
